\section{Type classes}
\topics{Type classes.}

These exercises are about type classes in Haskell. You can obtain the skeleton code by cloning the repository from GitHub:
\begin{minted}{bash}
$ git clone https://github.com/fpclass/lab-type-classes
\end{minted}

\taskLine 

\subsubsection{Monoids}

A \emph{monoid} is an algebraic structure which consists of a unit value and an associative, binary operation. In Haskell, we can declare a type class for types which are monoids:
\begin{minted}{haskell}
class Monoid a where
mempty  :: a
mappend :: a -> a -> a
mconcat :: [a] -> a
\end{minted}
Here, \texttt{mempty} represents the unit value and \texttt{mappend} represents the associative, binary operation. Instances of the \texttt{Monoid} type class should obey the following \emph{monoid laws}:
\begin{displaymath}
\begin{array}{lcrcl}
\textbf{Left identity} &\qquad & \mathit{mappend}~\mathit{mempty}~x & = & x \\
\textbf{Right identity} &\qquad & \mathit{mappend}~x~\mathit{mempty} & = & x \\
\textbf{Associativity} & \qquad & \mathit{mappend}~x~(\mathit{mappend}~y~z) & = & \mathit{mappend}~(\mathit{mappend}~x~y)~z \\
\textbf{Concatenation} & \qquad & \mathit{mconcat} & = & \mathit{foldr}~\mathit{mappend}~\mathit{mempty}
\end{array}
\end{displaymath}
We say that a type \emph{forms} a monoid if there is an instance of the \haskellIn{Monoid} type class for it which obeys the monoid laws. 

The \haskellIn{mconcat} function shown above is not necessary for a type to be a monoid, but it generalises the ordinary \haskellIn{concat} function on lists of lists and can easily be implemented with the help of \haskellIn{mappend} and \haskellIn{mempty}.

\taskLine 

\task{The \haskellIn{mconcat} function does nothing specific to any particular type. Specify a default implementation for the \haskellIn{mconcat} function in the declaration of the \haskellIn{Monoid} type class so that it obeys the fourth monoid law.}

\task{Does it matter whether you use \haskellIn{foldr} or \haskellIn{foldl} for the implementation of \haskellIn{mconcat}?}

\taskLine

\task{Implement \haskellIn{mempty} and \haskellIn{mappend} of the \haskellIn{Monoid} instance for \haskellIn{Int} so that they obey the monoid laws. Note that there are two possible implementations which satisfy the monoid laws -- can you think of both?}

\taskLine 

\task{Implement \haskellIn{mempty} and \haskellIn{mappend} of the \haskellIn{Monoid} instance for \texttt{\small [a]} so that they obey the monoid laws.}

\taskLine 

\task{Functions of type \texttt{\small a -> b} form a monoid if \texttt{\small b} is a monoid. Implement \haskellIn{mempty} and \haskellIn{mappend} of the \haskellIn{Monoid} instance for \texttt{\small a -> b} so that they obey the monoid laws. Note that there are no unit tests for this task as it would require tests for function equality.}

\taskLine 